\section*{Acknowledgements}

BDS would like to thank Michael Kuhlen for his work on the code in the
early stages of the project; the organizers of the AGORA project,
Ji-hoon Kim, Joel Primack, and Piero Madau for the original motivation
for the \texttt{Grackle} project; as well as Nick Gnedin, James
Wadsley, and Romeel Dav$\acute{\rm e}$ for providing useful
suggestions for additional functionality.  GLB acknowledges support
from NASA grant NNX15AB20G and NSF grant 1312888. SCOG acknowledges
support from the Deutsche Forschungsgemeinschaft via SFB 881, ``The
Milky Way System'' (sub-projects B1, B2 and B8) and SPP 1573,
``Physics of the Interstellar Medium'' (grant number GL 668/2-1), and
from the European Research Council under the European Community's
Seventh Framework Programme (FP7/2007-2013) via the ERC Advanced Grant
STARLIGHT (project number 339177).  NJG and MJT were supported by NSF
grant ACI-1535651 and by the Gordon and Betty Moore Foundation's
Data-Driven Discovery Initiative through Grant GBMF4651.  JHW is
supported by NSF grants AST-1333360 and AST-1614333 and Hubble theory
grants HST-AR-13895 and HST-AR-14326.  AE is supported by a NSF
Graduate Research Fellowship grant No. DGE-16-44869. We also thank the
NSF for computational resources provided through the XSEDE
program. JAR acknowledges support through the STFC capital grant
ST/L00075X/1 and the Marie Curie Research Fund, grant 699941.
BWO acknowledges support from  NASA through grants
NNX12AC98G, NNX15AP39G, and Hubble Theory Grants HST-AR-13261.01-A and
HST-AR-14315.001-A.  BWO was supported in part by the sabbatical
visitor program at the Michigan Institute for Research in Astrophysics
(MIRA) at the University of Michigan in Ann Arbor, and gratefully
acknowledges their hospitality.
Work by PA was performed in part under the auspices of
the U.S. Department of Energy by Lawrence Livermore National
Laboratiory under Contract DE-AC52-07NA27344.  \texttt{Grackle} was
also made possible by the open-source projects,
HDF\footnote{https://www.hdfgroup.org/},
h5py\footnote{http://www.h5py.org/}, matplotlib
\citep{2005ASPC..347...91B}, and NumPy \citep{numpy}.

