\section{Summary} \label{sec:summary}



\subsection{codes which have Grackle implementations}


\subsection{Future Directions} \label{Future_Directions}

\subsubsection{Future API changes} 
Replace the 18 array pointers to a single pointer to a struct of arrays?

\subsubsection{Including new rates and models in Grackle}
\jr{} The current code structure is highly integrated. This makes introducing new rates for the 
chemical network or cooling network a rather intricate task requiring multiple changes throughout the code. 
Apart from the fact that this is more time consuming it is also much more error prone. In a future release of the 
code the modularity of the code will be increased greatly. There will be a function to populate the species 
rate coefficients and a function to populate the cooling coefficients. Seperate template files can then be 
updated by a developer wishing to use their own rates. This file can then be included in the build and a flag
set to indicate the new rates be used in place of the old rates. Furthermore, a similar method will be 
implemented for solving the network. A template network solver will be available which the developer can use to 
implement a new network with a developer determined number of species. The developer will be responsible for
updating only three files to achieve a solution to their own chemical network. 

\subsubsection{Connection to Radiative Transfer}

\subsubsection{Multiple element cooling}

\subsubsection{Other heating sources?}
